---
title: "Take-home_EX02"
author: Edward
format: html
date: 4 Mar 2026
date-modified: 4 Mar 2026
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

# 1. Load Libraries

The following packages are used:

+----------------------------------+------------------------------------------------------------------------------------+
| Packages                         | Purpose                                                                            |
+==================================+====================================================================================+
| tidyverse.                       | Used for data manipulation, transformation, and visualisation.                     |
+----------------------------------+------------------------------------------------------------------------------------+
| cluster                          | Used for performing and evaluating cluster analysis.                               |
+----------------------------------+------------------------------------------------------------------------------------+
| GGally                           | Used for advanced exploratory data visualisations and variable relationship plots. |
+----------------------------------+------------------------------------------------------------------------------------+
| corrplot                         | Used for visualising correlation matrices.                                         |
+----------------------------------+------------------------------------------------------------------------------------+
| patchwork                        | Used for arranging multiple plots into a single layout.                            |
+----------------------------------+------------------------------------------------------------------------------------+
| knitr                            | Used for generating reproducible reports by integrating code, output, and text.    |
+----------------------------------+------------------------------------------------------------------------------------+

```{r}
pacman::p_load(tidyverse, cluster , GGally, corrplot, patchwork, knitr)
```

# 2. The Data

This analysis uses the **COFINFAD** dataset which is a comprehensive 2023 record of Colombian fintech activity.

The data set contains 2 files and only the **customer_data.csv** is used.

**K-means clustering** is applied to identify the distinct user profiles to optimize retention and product targeting.

<https://doi.org/10.1016/j.dib.2026.112484>

## 2.1 Load Data

```{r}
customer <- read_csv('Data/customer_data.csv')
skimr::skim(customer)
```

## 2.2 Data Prep

### 2.2.1 Feature selection & Scaling

-   The variables selected for K-means clustering were chosen to minimise redundancy by ensuring that no pair of variables exhibits a high correlation (greater than 0.8).

-   These four variables were retained because they provide a balanced representation of key customer behaviours that are likely to differentiate customer segments.

**Features selected:** monthly_transaction_count, avg_tx_value, active_products, customer_tenure

```{r}
#| code-fold: true
features <- c("monthly_transaction_count",
         "avg_tx_value",
         "active_products",
         "customer_tenure")

fin.data <- customer %>%
  select(all_of(features))

scaled_features <- scale(fin.data)

fin.cor <- cor(fin.data, use = "complete.obs")

corrplot(
  fin.cor,
  method = "ellipse",
  order = "AOE",
  col = colorRampPalette(c("blue", "white", "red"))(200),
  tl.col = "black"
)
```

```         
```

# 3 The Model

## 3.1 Feeding the model

### 3.1.1 Running the Elbow Plot

The elbow plot is run to allow use to narrow down to the number of cluster that potentially give us the better separation.

```{r}
#| code-fold: true
# set up no. of max clusters to run
n = 10
# Create empty list
wcss <- numeric(n)

# Create loop and extract SSW into list
set.seed(888)
for (k in 1:n) {
  km <- kmeans(scaled_features, centers = k, nstart = 10,
                       iter.max = 200, algorithm = "Lloyd")
  wcss[k] <- km$tot.withinss
}

plot(1:n, wcss, type = "b",
     xlab = "Number of Clusters",
     ylab = "Within Cluster SSE",
     main = "Elbow Plot of K-Means Clustering")
```

### 3.1.2 Running the Kmeans model

-   From the elbow plot the ideal number of clusters are set to be between 3-5.

-   Although k = 5 produced a slightly higher silhouette score, the fifth cluster contained only 24 observations, indicating poor stability. Therefore **k = 4** was selected as it provides a **balance between cluster quality and interpretability**.

-   To allow for better explainability, PCA is not applied in the model.

-   We are using Lloyd algorithm as it is more stable for larger data-set.

```{r}
#| code-fold: true
# Run k means based on n clusters
kmeans_model <- kmeans(scaled_features, centers = 4, nstart = 10, iter.max = 200, algorithm = "Lloyd")

# Assign labels back to original data
customer$cluster <- factor(kmeans_model$cluster)

# Assign cluster to count size to prevent shifting
cluster_order <- customer %>%
  count(cluster) %>%
  arrange(desc(n)) %>%
  mutate(new_cluster = row_number())

customer <- customer %>%
  left_join(cluster_order %>% select(cluster, new_cluster),
            by = "cluster") %>%
  mutate(cluster = factor(new_cluster)) %>%
  select(-new_cluster)
```

# 4 Performance Evaluation

**Overall Performance Summary:**

+--------------+--------------+--------------------+
| K Clusters   | Silhouette   | Variance Explained |
+==============+==============+====================+
| 3            | 0.3907       | 0.36               |
+--------------+--------------+--------------------+
| 4            | 0.4272       | 0.49               |
+--------------+--------------+--------------------+
| 5\*          | 0.4308       | 0.64               |
+--------------+--------------+--------------------+

\*Cluster of 5 is not selected as the 5th cluster is only with 24 members which is too small.

## 4.1 Silhouette

Silhouette measures **how well each observation fits within its assigned cluster as compare to other clusters**.

Silhouette value is between **-1 and 1**.

• **Close to 1** → well separated\
• **Around 0** → overlapping clusters\
• **Negative** → likely misclassified

```{r}
#| code-fold: true
sil <- silhouette(kmeans_model$cluster,
                  dist(scaled_features))

cat("Silouette:", mean(sil[,"sil_width"]))
```

## 4.2 Variance Explained

-   The proportion of variance explained is computed as the **ratio of between-cluster sum of squares to total sum of squares**.

-   Variance explained measures the proportion of **total variation in the dataset that is captured** by the clustering structure.

-   The v**alue ranges from 0 to 1**, where higher values indicate greater separation between clusters.

**Clusters Distributions**

```{r}
#| code-fold: true
# Calculate number of observations in each cluster.
cluster_size <- table(customer$cluster)
cluster_size

```

```{r}
#| code-fold: true
# Calculate between-cluster Sum of Squares
btwn_clus <- kmeans_model$betweenss

# Calculate total Sum of Squares
tss <- kmeans_model$totss

# Proportion of total variance explained by clustering
var_exp <- (kmeans_model$betweenss / kmeans_model$totss)

cat(sprintf("Between-cluster SS: %.2f\n", btwn_clus))
cat(sprintf("Total SE: %.2f\n", tss))
cat(sprintf("Proportion of variance explained: %.2f\n", var_exp))
```

# 5 The Clusters

## 5.1 Cluster size

4 clusters were selected as:

-   It is able to explain the variance as compared to 5 clusters

-   The feature are more differentiated

```{r}
#| code-fold: true
cluster_size <- customer %>%
  count(cluster) %>%
  mutate(pct = n / sum(n))

p1 <- ggplot(cluster_size,
       aes(x = cluster, y = n, fill = cluster)) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  geom_text(
    aes(label = scales::percent(pct, accuracy = 0.1)),
    vjust = -0.3
  ) +
  theme_minimal() +
  labs(title = "Number of Customers by Cluster",
       x = "Cluster",
       y = "Customers")

p1
```

## 5.2 Cluster Characteristics

### 5.2.1 Parallel plot

```{r}
#| code-fold: true
# Group of feature mean by clusters
cluster_mean <- customer %>%
  group_by(cluster) %>%
  summarise(across(c(all_of(features)),
                   mean))

# Plot parallel plot
p2 <- ggparcoord(cluster_mean,
           columns = 2:ncol(cluster_mean),
           groupColumn = "cluster",
           scale = "uniminmax") +
  geom_line(linewidth = 1.5) +
  geom_point(size = 2) + 
  labs(title = "Parallel plot by clusters mean") +
  scale_color_brewer(palette = "Set2", guide = guide_legend(reverse = FALSE))+
  theme_minimal()


p2
```

### **5.2.2 Boxplot of Selected features**

```{r, fig.width= 12}
#| code-fold: true
customer_piv <- customer %>%
  select(cluster,
         all_of(features)) %>%
  pivot_longer(-cluster,
               names_to = "feature",
               values_to = "value")

p3 <- ggplot(customer_piv,
       aes(x = cluster,
           y = value,
           fill = cluster)) +
  geom_boxplot() +
  facet_wrap(~ feature,
             scales = "free_y") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(title = "Cluster Profiles Across Features",
       x = "Cluster",
       y = "Value")

p3
```

### **5.2.3 Median by Cluster Group**

```{r}
#| code-fold: true
cluster_median <- customer %>%
  group_by(cluster) %>%
  summarise(across(all_of(features), median))
            
cluster_median %>%
  kable(caption = "Median Feature Values by Cluster")

```

### 5.2.4 Detailed Cluster Comparison

The clusters can be differentiated by the characteristics below:

+-------------+----------------------------------------------------+----------------------------------------------------------------------+
| Cluster     | Segment Type                                       | Characteristics                                                      |
+=============+====================================================+======================================================================+
| 1           | Active Low Value and Long Term Customer            | -   Account for the largest pool of customers (54.4%)                |
|             |                                                    |                                                                      |
|             |                                                    | -   Highest monthly transaction count (Median = 2)                   |
|             |                                                    |                                                                      |
|             |                                                    | -   Customer with the highest tenure (Median = 11.7)                 |
|             |                                                    |                                                                      |
|             |                                                    | -   Low number of active products (Median = 1)                       |
|             |                                                    |                                                                      |
|             |                                                    | -   Lowest average transaction (Median = 1675191)                    |
+-------------+----------------------------------------------------+----------------------------------------------------------------------+
| 2           | Multi-product Low Value and Long Term Customer     | -   Customer with the highest tenure (Median = 11.7)                 |
|             |                                                    |                                                                      |
|             |                                                    | -   Very high monthly transaction count (Median = 2.09)              |
|             |                                                    |                                                                      |
|             |                                                    | -   Very small average transaction amount (similar to cluster 1)     |
|             |                                                    |                                                                      |
|             |                                                    | -   Hold the highest number of active products (Median = 4)          |
+-------------+----------------------------------------------------+----------------------------------------------------------------------+
| 3           | New and Low Value Customer                         | -   Customer with lowest tenure (Median = 10.07)                     |
|             |                                                    |                                                                      |
|             |                                                    | -   Lowest transaction count (Median = 1.67)                         |
|             |                                                    |                                                                      |
|             |                                                    | -   Average transaction value is relatively low (Median = 1,725,881) |
+-------------+----------------------------------------------------+----------------------------------------------------------------------+
| 4           | Premium Customer: Moderately Active and High Value | -   Customer nearing the highest tenure (Median = 11.63)             |
|             |                                                    |                                                                      |
|             |                                                    | -   Very high monthly transaction count (Median = 2)                 |
|             |                                                    |                                                                      |
|             |                                                    | -   Average transaction amount is the highest (Median = 13,719,926)  |
+-------------+----------------------------------------------------+----------------------------------------------------------------------+

## 5.3 Clusters Summary

The Key features for the cluster can be summarised by the Heatmap:

```{r}
#| code-fold: true
# Normalise fetaure to scale of 0 to 1
cluster_normalised <- cluster_median %>%
  mutate(across(-cluster, ~ (. - min(.)) / (max(.) - min(.))))

# Make tabular format
cluster_long <- cluster_normalised %>%
  pivot_longer(-cluster,
               names_to = "feature",
               values_to = "value")


ggplot(cluster_long,
       aes(x = feature,
           y = cluster,
           fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradientn(
    colours = c("#e5f5f9", "#99d8c9", "#2ca25f")
  )+
  theme_minimal() +
  labs(title = "Cluster Feature Heatmap",
       x = "Customer Behaviour Features",
       y = "Cluster",
       fill = "Relative Intensity")
```

The 4 clusters are summarized as below:

-   **Cluster 1**: Active Low Value and Long Term Customer

-   **Cluster 2**: Multi-product Low Value and Long Term Customer

-   **Cluster 3**: New and Low Value Customer

-   **Cluster 4**: Premium Customer: Moderately Active and High Value

## References:

*Kam, Tin Seong. (2025).6  Visual Correlation Analysis – R for Visual Analytics*. (2023, December 4). Netlify.app. <https://r4va.netlify.app/chap06‌>

*Kam, Tin Seong. (2025).15  Visual Multivariate Analysis with Parallel Coordinates Plot – R for Visual Analytics*. (2023, December 4). Netlify.app. <https://r4va.netlify.app/chap15>

*Kam, Tin Seong. (2025).16  Treemap Visualisation with R – R for Visual Analytics*. (2023, December 4). Netlify.app. <https://r4va.netlify.app/chap16>

‌*Heat map in ggplot2*. (2021, November 2). R CHARTS \| a Collection of Charts and Graphs Made with the R Programming Language. <https://r-charts.com/correlation/heat-map-ggplot2/>

‌

‌
