---
title: "Hands-on_Ex 04d"
author: "Edward"
---

# **8 Funnel Plots for Fair Comparisons**

## **8.2 Installing and Launching R Packages**

In this exercise, four R packages will be used. They are:

-   **FunnelPlotR** for creating funnel plot.

-   **ggplot2** for creating funnel plot manually.

-   **knitr** for building static html table.

-   **plotly** for creating interactive funnel plot.

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

## **8.3 Importing Data**

```{r}
#| message: false
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

## **8.4 FunnelPlotR methods**

[**FunnelPlotR**](https://nhs-r-community.github.io/FunnelPlotR/) package uses ggplot to generate funnel plots. It requires a `numerator` (events of interest), `denominator` (population to be considered) and `group`. The key arguments selected for customisation are:

-   `limit`: plot limits (95 or 99).

-   `label_outliers`: to label outliers (true or false).

-   `Poisson_limits`: to add Poisson limits to the plot.

-   `OD_adjust`: to add overdispersed limits to the plot.

-   `xrange` and `yrange`: to specify the range to display for axes, acts like a zoom function.

-   Other aesthetic components such as graph title, axis labels etc.

#### 8.4.1 FunnelPlotR methods: The basic plot

```{r}
funnel_plot(
  .data = covid19,
  numerator = Positive,
  denominator = Death,
  group = `Sub-district`,
  title = "The basic plot"
)
```

#### 8.4.2 FunnelPlotR methods: Makeover 1

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",     #<<
  x_range = c(0, 6500),  #<<
  y_range = c(0, 0.05),   #<<
  title = "The basic plot M1"
)
```

#### 8.4.3 FunnelPlotR methods: Makeover 2

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  x_range = c(0, 6500),  
  y_range = c(0, 0.05),
  label = NA,
  title = stringr::str_wrap("Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases",width=50), #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)
```

## **8.5 Funnel Plot for Fair Visual Comparison: ggplot2 methods**

#### **8.5.1 Computing the basic derived fields**

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
```

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

#### 8.5.2 Calculate lower and upper limits for 95% and 99.9% CI

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

#### 8.5.3 Plotting a static funnel plot

```{r}
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            linewidth = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            linewidth = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            linewidth = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            linewidth = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             linewidth = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))

```

#### **8.5.4 Interactive Funnel Plot: plotly + ggplot2**

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```
