# **21  Choropleth Mapping with R**

---
title: "Hands-on Exercise 8a"
author: "Edward"
---

## **21.2 Getting Started**

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## **21.3 Importing Data into R**

```{r}
#| message: false
mpsz <- st_read(dsn = "Data/Master Plan 2014 Subzone Boundary (Web) (SHP)", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
#| message: false
mpsz
```

### **21.3.3 Importing Attribute Data into R**

Import *respopagsex2011to2020.csv* file into RStudio and save the file into an R dataframe called *popagsex*

```{r}
#| message: false
popdata <- read_csv("Data/respopagesextod2011to2020.csv")
```

### **21.3.4 Data Preparation**

#### 21.3.4.1 Data wrangling

```{r}
#| message: false
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

#### 21.3.4.2 Joining the attribute data and geospatial data

Convert the values in PA and SZ fields to uppercase

```{r}
popdata2020 <- popdata2020 %>%
  mutate(across(c(PA, SZ), toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "Data/mpszpop2020.rds")
```

## **21.4 Choropleth Mapping Geospatial Data Using *tmap***

### **21.4.1 Plotting a choropleth map quickly by using *qtm()***

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### **21.4.2 Creating a choropleth map by using *tmap*’s elements**

**qtm( )advantages:** Able to draw a choropleth map quickly and easily,

**Disadvantge of *qtm():*** It makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
#| message: false
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

#### 21.4.2.1 Drawing a base map

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 21.4.2.2 Drawing a choropleth map using *tm_polygons()*

With line border

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### 21.4.2.3 Drawing a choropleth map using *tm_fill()* and \*tm_border()\*\*

Without line border

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY") +
  tm_borders(lwd = 0.01,  
             fill_alpha = 0.1)
```

The *alpha* argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).

Beside *alpha* argument, there are three other arguments for *tm_borders()*, they are:

-   *col* = border colour,

-   *lwd* = border line width. The default is 1, and

-   *lty* = border line type. The default is “solid”.

### **21.4.3 Data classification methods of tmap**

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()* or *tm_polygons()* will be used.

#### 21.4.3.1 Plotting choropleth maps with built-in classification methods

**Style=jenks**

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

**Style=equal**

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "equal",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

#### 21.4.3.2 Plotting choropleth map with customer break

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
#| message: false
#| warning: false
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(fill_alpha = 0.5)
```

### **21.4.4 Colour Scheme**

#### 21.4.4.1 Using ColourBrewer palette

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

### **21.4.5 Map Layouts**

#### 21.4.5.1 Map Legend

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5,
        values = "brewer.greens"),
      fill.legend = tm_legend(
        title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```

#### 21.4.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

```{r}
#| message: false
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### 21.4.5.3 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

In the code chunk below, `tm_compass()`, `tm_scale_bar()` and `tm_grid()` are used to add compass, scale bar and grid lines onto the choropleth map.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

**To reset to default stype**

```{r}
#| message: false
tmap_style("white")
```

#### **21.4.6 Drawing Small Multiple Choropleth Maps**

#### 21.4.6.1 By assigning multiple values to at least one of the aesthetic arguments

```{r}
#| message: false
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues",
          ncols = 1) +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

```{r}
#| message: false
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

```{r}
#| message: false
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### 21.4.6.2 By defining a group-by variable in *tm_facets()*

Multiple small choropleth maps are created by using `tm_facets()`.

```{r}
#| message: false
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### 21.4.6.3 By creating multiple stand-alone maps with `tmap_arrange()`

```{r}
#| message: false
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### **21.4.7 Mappping Spatial Object Meeting a Selection Criterion**

```{r}
#| message: false
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY",
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 2,
            legend.position = c("right", "bottom"),
            legend.outside = TRUE,
            legend.outside.position = "right",
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
