# **22  Visualising Geospatial Point Data**

---
title: "Hands-on_Ex08b"
author: "Edward"
---

# **24 Getting Started**

```{r}
#| message: false
pacman::p_load(sf, tmap, tidyverse)
```

# **25 Geospatial Data Wrangling**

## **25.1 The data**

The data set use for this hands-on exercise is called *SGPools_svy21*. The data is in csv file format.

Figure below shows the first 15 records of SGPools_svy21.csv. It consists of seven columns. The XCOORD and YCOORD columns are the x-coordinates and y-coordinates of SingPools outlets and branches. They are in [Singapore SVY21 Projected Coordinates System](https://www.sla.gov.sg/sirent/CoordinateSystems.aspx).

## **25.2 Data Import and Preparation**

```{r}
sgpools <- read_csv("Data/SGPools_svy21.csv")
```

```{r}
list(sgpools)
```

## **25.3 Creating a sf data frame from an aspatial data frame**

The code chunk below converts sgpools data frame into a simple feature data frame by using *st_as_sf()* of **sf** packages

-   The *crs* argument required you to provide the coordinates system in epsg format. [EPSG: 3414](https://epsg.io/3414) is Singapore SVY21 Projected Coordinate System. We can search for other country’s epsg code by refering to [epsg.io](https://epsg.io/).

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

```{r}
#| message: false
list(sgpools_sf)
```

# **26 Drawing Proportional Symbol Map**

```{r}
#| message: false
tmap_mode("view")
```

## **26.1 It all started with an interactive point symbol map**

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
           size = 1,
           col = "black",
           lwd = 1)
```

## **26.2 Lets make it proportional**

To draw a proportional symbol map, we need to assign a numerical variable to the size visual attribute.

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## **26.3 Lets give it a different colour**

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## **26.4 I have a twin brothers :)**

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1) + 
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

To switch **tmap**’s Viewer back to plot mode by using the code chunk below.

```{r}
#| message: false
tmap_mode("plot")
```
