---
title: "Hands-on Exercise 7a"
author: "Edward"
---

## **17.2 Getting Started**

## **17.3 Import library**

```{r}
pacman::p_load(scales,viridis,lubridate,ggthemes,gridExtra,readxl,knitr,data.table,tidyverse)
```

```{r}
library(CGPfunctions,ggHoriPlot)
```

## **17.4 Plotting Calendar Heatmap**

### **17.4.2 Importing the data**

```{r}
attacks <- read.csv('Data/eventlog.csv')
```

### **17.4.3 Examining the data structure**

```{r}
kable(head(attacks))
```

### **17.4.4 Data Preparation**

Step 1: Deriving *weekday* and *hour of day* fields

```{r}
make_hr_wkday <- function(ts,sc,tz){
  real_times <- ymd_hms(ts,
                        #uses first timezone in TZ
                        tz = tz[1],
                        quiet = TRUE)
  
  dt <- data.table(source_country=sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
}
```

Step 2: Deriving the attacks tibble data frame

```{r}
wkday_levels <- c('Saturday','Friday','Thursday',
                  'Wednesday','Tuesday','Monday','Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp,
                   .$source_country,
                   .$tz)) %>%
  ungroup() %>%
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour = factor(
      hour, levels = 0:23))
```

```{r}
kable(head(attacks))
```

### **17.4.5 Building the Calendar Heatmaps**

```{r}
#| warnings: false
grouped <- attacks %>%
  count(wkday, hour) %>%
  ungroup() %>%
  na.omit()

ggplot(grouped,
       aes(hour,
           wkday,
           fill = n)) +
  geom_tile(color = 'white',
             linewidth = 0.1) +
  theme_tufte() +
  coord_equal() +
  scale_fill_gradient(name = '# of attacks',
                      low = 'sky blue',
                      high = 'dark blue') +
  
  labs(x = NULL,
       y = NULL,
       title = 'Attacks by Weekday and time of day') +
  
  theme(axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6))
  

```

### **17.4.6 Building Multiple Calendar Heatmaps**

**Challenge:** Building multiple heatmaps for the top four countries with the highest number of attacks.

### **17.4.7 Plotting Multiple Calendar Heatmaps**

Step 1: Deriving attack by country object

In order to identify the top 4 countries with the highest number of attacks, The following was done:

-   count the number of attacks by country,

-   calculate the percent of attackes by country, and

-   save the results in a tibble data frame.

```{r}
attacks_by_country <- 
  count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```

Step 2: Preparing the tidy data frame

Extract the attack records of the top 4 countries from *attacks* data frame and save the data in a new tibble data frame (i.e. *top4_attacks*).

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country,wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
  
```

### **17.4.8 Plotting Multiple Calendar Heatmaps**

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r}
#| warning: false
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          linewidth = 0.1) + 
  theme_tufte() + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## **17.5 Plotting Cycle Plot**

### **17.5.1 Step 1: Data Import**

```{r}
air <- read_excel('Data/arrivals_by_air.xlsx')
```

### **17.5.2 Step 2: Deriving month and year fields**

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### **17.5.3 Step 4: Extracting the target country**

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### **17.5.4 Step 5: Computing year average arrivals by month**

`group_by()` and `summarise()` of **dplyr** compute year average arrivals by month.

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(Vietnam))
```

### **17.5.5 Srep 6: Plotting the cycle plot**

```{r}
str(Vietnam)
str(hline.data)
```

```{r, fig.width= 20, fig.height= 10}
#| warning: false
ggplot(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`)) + 
  geom_line(colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linewidth=0.5, 
             colour="red") + 
  facet_grid(~month) +
  scale_x_continuous(breaks = seq(2010, 2019, 2),expand = c(0,0)) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme(
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20)
) +
  theme_tufte() +
  theme_minimal()+
theme(
  axis.text.x = element_text(size = 15, angle=45),
  axis.text.y = element_text(size = 20),
  strip.text = element_text(size = 20),
  plot.title = element_text(size = 30),
  panel.spacing = unit(1, "lines"),
  panel.background = element_rect(fill = "grey95"),
  strip.background = element_rect(
    fill = "lightblue",
    colour = "grey95"
  )
)
```

## **17.6 Plotting Slopegraph**

```{r}
#| message: false
rice <- read_csv("Data/rice.csv")
```

### **17.6.2 Step 2: Plotting the slopegraph**

```{r}
#| warning: false
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Rice yield have increased for all countries \nwith China seeing the most growth.")
```
