# **14  Heatmap for Visualising and Analysing Multivariate Data**

---
title: "Hands-on_EX04c"
author: "Edward"
---

## **14.2 Installing and Launching R Packages**

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## **14.3 Importing and Preparing The Data Set**

#### 14.3.1 Importing the data set

```{r}
#| message: false
wh <- read_csv("data/WHData-2018.csv")
```

#### 14.3.2 Preparing the data

This command assigns country names as row identifiers in the dataframe, replacing the default numeric row indices.

```{r}
row.names(wh) <- wh$Country
```

#### 14.3.3 Transforming the data frame into a matrix

The data was loaded into a data frame, but it has to be a data matrix to make your heatmap.

The code chunk below will be used to transform *wh* data frame into a data matrix.

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## **14.4 Static Heatmap**

-   [heatmap()](https://www.rdocumentation.org/packages/stats/versions/3.6.0/topics/heatmap)of R stats package. It draws a simple heatmap.

-   [heatmap.2()](https://www.rdocumentation.org/packages/gplots/versions/3.0.1.1/topics/heatmap.2) of **gplots** R package. It draws an enhanced heatmap compared to the R base function.

-   [pheatmap()](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12/topics/pheatmap) of [**pheatmap**](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12) R package. **pheatmap** package also known as Pretty Heatmap. The package provides functions to draws pretty heatmaps and provides more control to change the appearance of heatmaps.

-   [**ComplexHeatmap**](https://bioconductor.org/packages/release/bioc/html/ComplexHeatmap.html) package of R/Bioconductor package. The package draws, annotates and arranges complex heatmaps (very useful for genomic data analysis). The full reference guide of the package is available [here](https://jokergoo.github.io/ComplexHeatmap-reference/book/).

-   [**superheat**](https://cran.r-project.org/web/packages/superheat/) package: A Graphical Tool for Exploring Complex Datasets Using Heatmaps. A system for generating extendable and customizable heatmaps for exploring complex datasets, including big data and data with multiple data types. The full reference guide of the package is available [here](https://rlbarter.github.io/superheat/).

#### 14.4.1 heatmap() of R Stats

-   By default, **heatmap()** plots a cluster heatmap. The arguments ***Rowv=NA*** and ***Colv=NA*** are used to switch off the option of plotting the row and column dendrograms.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

Red cells denotes small values, and red small ones. The Happiness Score variable have relatively higher values, what makes that the other variables with small values all look the same.

Thus, we need to normalize this matrix. This is done using the ***scale*** **argument**. It can be applied to rows or to columns.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

## **14.5 Creating Interactive Heatmap**

#### 14.5.1 Working with heatmaply

```{r}
#| warning: false
heatmaply(mtcars)
```

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

## **14.5.2 Data trasformation**

#### 14.5.2.1 Scaling method

When all variables are came from or assumed to come from some normal distribution, then scaling (i.e.: subtract the mean and divide by the standard deviation) would bring them all close to the standard normal distribution.

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

#### 14.5.2.2 Normalising method

When variables in the data comes from possibly different (and non-normal) distributions.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

#### 14.5.2.3 Percentising method

This is similar to ranking the variables, but instead of keeping the rank values, divide them by the maximal rank.

This is done by using the ecdf of the variables on their own values, bringing each value to its empirical percentile.

The benefit of the percentize function is that each value has a relatively clear interpretation, it is the percent of observations that got that value or below it.

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

## **14.5.3 Clustering algorithm**

**heatmaply** supports a variety of hierarchical clustering algorithm. The main arguments provided are:

-   ***distfun***: function used to compute the distance (dissimilarity) between both rows and columns. Defaults to dist. The options “pearson”, “spearman” and “kendall” can be used to use correlation-based clustering, which uses as.dist(1 - cor(t(x))) as the distance metric (using the specified correlation method).

-   ***hclustfun***: function used to compute the hierarchical clustering when *Rowv* or *Colv* are not dendrograms. Defaults to *hclust*.

-   ***dist_method*** default is NULL, which results in “euclidean” to be used. It can accept alternative character strings indicating the method to be passed to distfun. By default *distfun* is “dist”” hence this can be one of “euclidean”, “maximum”, “manhattan”, “canberra”, “binary” or “minkowski”.

-   ***hclust_method*** default is NULL, which results in “complete” method to be used. It can accept alternative character strings indicating the method to be passed to *hclustfun*. By default hclustfun is hclust hence this can be one of “ward.D”, “ward.D2”, “single”, “complete”, “average” (= UPGMA), “mcquitty” (= WPGMA), “median” (= WPGMC) or “centroid” (= UPGMC).

In general, a clustering model can be calibrated either manually or statistically.

## **14.5.4 Manual approach**

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

## **14.5.5 Statistical approach**

First, the *dend_expend()* will be used to determine the recommended clustering method to be used.

We select the higher optimum value.

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

Figure below shows that k=3 would be good.

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

## **14.5.6 Seriation**

**Optimal Leaf Ordering (OLO)**. This algorithm starts with the output of an agglomerative clustering algorithm and produces a unique ordering, one that flips the various branches of the dendrogram around so as to minimize the sum of dissimilarities between adjacent leaves. Here is the result of applying Optimal Leaf Ordering to the same clustering result as the heatmap above.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

Another option is “GW” (Gruvaeus and Wainer) which aims for the same goal but uses a potentially faster heuristic.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

The option “mean” gives the output we would get by default from heatmap functions in other packages such as gplots::heatmap.2.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

## **14.5.7 Working with colour palettes**

The default colour palette uses by **heatmaply** is *viridis*. heatmaply users, however, can use other colour palettes in order to improve the aestheticness and visual friendliness of the heatmap.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

## **14.5.8 The finishing touch**

In the code chunk below the following arguments are used:

-   *k_row* is used to produce 5 groups.

-   *margins* is used to change the top margin to 60 and row margin to 200.

-   *fontsizw_row* and *fontsize_col* are used to change the font size for row and column labels to 4.

-   *main* is used to write the main title of the plot.

-   *xlab* and *ylab* are used to write the x-axis and y-axis labels respectively.

    ```{r,fig.height = 12}
    heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
              Colv=NA,
              seriate = "none",
              colors = Blues,
              k_row = 5,
              margins = c(NA,200,60,NA),
              fontsize_row = 4,
              fontsize_col = 5,
              main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
              xlab = "World Happiness Indicators",
              ylab = "World Countries"
              )
    ```
