---
title: "Hands-on_EX05e"
author: "Edward"
---

# **16  Treemap Visualisation with R**

## 16.2 Installing and Launching R Packages

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

## **16.3 Data Wrangling**

#### 16.3.1 Importing the data set

```{r}
#| message: false
realis2018 <- read_csv("data/realis2018.csv")
```

#### 16.3.2 Data Wrangling and Manipulation

In this section, we will perform the following steps to manipulate and prepare a data.frtame that is appropriate for treemap visualisation:

-   group transaction records by *Project Name*, *Planning Region*, *Planning Area*, *Property Type* and *Type of Sale*, and

-   compute *Total Unit Sold*, *Total Area*, *Median Unit Price* and *Median Transacted Price* by applying appropriate summary statistics on *No. of Units*, *Area (sqm)*, *Unit Price (\$ psm)* and *Transacted Price (\$)* respectively.

Grouping \[*group_by()* and *summarize()*\] affects the verbs as follows:

-   grouped *select()* is the same as ungrouped *select()*, except that grouping variables are always retained.

-   grouped *arrange()* is the same as ungrouped; unless you set *.by_group = TRUE*, in which case it orders first by the grouping variables.

-   ***mutate()*** **and *filter()*** are most useful in conjunction with window functions (like *rank()*, or *min(x) == x*). They are described in detail in vignette(“window-functions”).

-   *sample_n()* and *sample_frac()* sample the specified number/fraction of rows in each group.

-   *summarise()* computes the summary for each group.

#### 16.3.3 Grouped summaries without the Pipe

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

Aggregation functions such as *sum()* and *meadian()* obey the usual rule of missing values: if there’s any missing value in the input, the output will be a missing value. The argument *na.rm = TRUE* removes the missing values prior to computation

## **16.4 Designing Treemap with treemap Package**

#### 16.4.1 Designing a static treemap

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

#### 16.4.2 Using the basic arguments

-   **index**

    -   The index vector must consist of at least two column names or else no hierarchy treemap will be plotted.

    -   If multiple column names are provided, such as the code chunk above, the first name is the highest aggregation level, the second name the second highest aggregation level, and so on.

-   **vSize**

    -   The column must not contain negative values. This is because it’s vaues will be used to map the sizes of the rectangles of the treemaps.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

#### 16.4.3 Working with *vColor* and *type* arguments

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

#### 16.4.4 Colours in treemap package

The “**value**” treemap considers palette to be a diverging color palette (say ColorBrewer’s “RdYlBu”), and maps it in such a way that 0 corresponds to the middle color (typically white or yellow), -max(abs(values)) to the left-end color, and max(abs(values)), to the right-end color. 

The “**manual**” treemap simply maps min(values) to the left-end color, max(values) to the right-end color, and mean(range(values)) to the middle color.

#### **16.4.5 The “value” type treemap**

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

#### 16.4.6 The “manual” type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

It is not wise to use diverging colour palette such as *RdYlBu* if the values are all positive or negative

Thus a single colour palette such as Blues should be used

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

#### 16.4.7 Treemap Layout

*treemap()* supports two popular treemap layouts, namely: “[squarified]{.underline}” and “[pivotSize]{.underline}”. The default is “pivotSize”.

The **squarified treemap** algorithm (Bruls et al., 2000) produces good aspect ratios, but ignores the sorting order of the rectangles (sortID).

The ordered treemap, **pivot-by-size**, algorithm (Bederson et al., 2002) **takes the sorting order (sortID) into account** while aspect ratios are still acceptable.

#### 16.4.8 Working with *algorithm* argument

**Squarified**

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

**Pivot Size**

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## **16.5 Designing Treemap using treemapify Package**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```

#### 16.5.2 Defining hierarchy

**Group by Planning Region**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

**Group by Planning Area**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap()
```

**Adding boundary line**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

## **16.6 Designing Interactive Treemap using d3treeR**

#### 16.6.1 Installing *d3treeR* package

1.  If this is the first time you install a package from github, you should install *devtools* package by using the code below or else you can skip this step.

**install.packages("devtools")**

2.  Next, you will load the *devtools* library and install the package found in github by using the codes below.

```{r}
library(devtools)
install_github("timelyportfolio/d3treeR")
```

3.  Now you are ready to launch *d3treeR* package

```{r}
library(d3treeR)
```

#### 16.6.2 Designing An Interactive Treemap

1.  ***treemap()*** is used to build a treemap by using selected variables in condominium data.frame. The treemap created is save as object called ***tm***.

```{r}
tm <- treemap(realis2018_summarised,
        index=c("Planning Region", "Planning Area"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        title="Private Residential Property Sold, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

2.  Then *d3tree()* is used to build an interactive treemap.

```{r}
d3tree(tm,rootname = "Singapore" )
```
